---
title: "Assignment 1"
author: "Don Padmaperuma (Geeth)"
date: "1/27/2020"
output: 
  html_document:
    toc: True
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(pracma)
```


## 1. Problem set 1

### (1) Calculate the dot product u.v where u = [0.5;0.5] and v = [3;-4]


```{r}

u <- c(0.5, 0.5)
v <- c(3, -4)

dot_uv <- dot(u, v)
dot_uv
```

### (2) What are the lengths of u and v? Please note that the mathematical notion of the length of a vector is not the same as a computer science definition.

```{r}
len_u <- sqrt(u[1]^2+u[2]^2)
len_v <- sqrt(v[1]^2+v[2]^2)
len_u;
len_v
```

### (3) What is the linear combination: 3u − 2v?

```{r}
3*u - 2*v
```

### (4) What is the angle between u and v

```{r}
cos_theta <- dot_uv/(len_u*len_v)
cos_theta
```
```{r}
angle_theta <- (180/pi)*acos(cos_theta)
angle_theta
```

## 2. Problem set 2

Set up a system of equations with 3 variables and 3 constraints and solve for x. Please
write a function in R that will take two variables (matrix A & constraint vector b) and
solve using elimination. Your function should produce the right answer for the system of
equations for any 3-variable, 3-equation system. You don’t have to worry about degenerate
cases and can safely assume that the function will only be tested with a system of equations
that has a solution. Please note that you do have to worry about zero pivots, though. Please
note that you should not use the built-in function solve to solve this system or use matrix
inverses. The approach that you should employ is to construct an Upper Triangular Matrix
and then back-substitute to get the solution. Alternatively, you can augment the matrix
A with vector b and jointly apply the Gauss Jordan elimination procedure.


```{r}
# Creating a matrix A and constrant b
A <- matrix(c(1, 2, -1, 1, -1, -2, 3, 5, 4), nrow=3, ncol=3) 
b <- matrix(c(1, 2, 6), nrow=3, ncol=1)

A;
b
```

Function for Guassian Elimination   
Reference:   
[https://stackoverflow.com/questions/16044377/how-to-do-gaussian-elimination-in-r-do-not-use-solve]
[https://martin-thoma.com/solving-linear-equations-with-gaussian-elimination/]

```{r}
x = function(A, b){
  r <- dim(A)[1]
  c <- dim(A)[2]+dim(b)[2]
  UT <- matrix(c(A, b), nrow=r, ncol=c)
  for (j in 1:(c-2)) {
    for (i in (j+1):r) {
      UT[i,] <- UT[i,]-UT[j,]*UT[i,j]/UT[j,j]  
    }
  }
  UT[r,] <- UT[r,]/UT[r,r]
  xn <- numeric(r)
  xn[r] = UT[r,c]
  for (k in (r-1):1) {
    t = 0
    for (m in (k+1):r) {
      s = UT[k,m]*xn[m]
      t = t + s
    }
    xn[k] = (UT[k,c] - t) / UT[k,k]
  }
  x <- round(xn,2)
return(x)
}
```


test the function with the matrix A and constraint vector b above and it should produce a solution x = [−1.55, −0.32, 0.95]

```{r}
x(A,b)
```

